### 2021-05-29 metric_analysis.Rmd
Update metric_analysis.Rmd

---
title: "metric analysis"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
    toc_float: yes
    number_sections: true
    code_folding: hide
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)

library(readr)
resist <- read_csv("ED-outputs/results/metric/resistance_values.csv")
met <- read_csv("ED-outputs/results/constant_annual_met_data.csv")
```

# Met data

```{r, echo=FALSE}
met %>% 
  # hgt is constant; remove
  filter(variable != "hgt") %>%   
  group_by(met, variable) %>% 
  summarise(value = mean(value), .groups = "drop") ->
  met2

library(ggplot2)
theme_set(theme_minimal())
ggplot(met2, aes(met, value)) +
  geom_point() + 
  facet_wrap(~variable, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 90))

library(tidyr)
met2_wide <- pivot_wider(met2, id_cols = "met", names_from = "variable")
pca_met <- prcomp(met2_wide[-1], scale. = TRUE)
biplot(pca_met, scale = 0)
pairs(met2_wide[-1])
```

# Metric data

```{r}
ggplot(resist, aes(trough_resistance, fill = severity)) +
  geom_histogram(bins = 20) + 
  facet_wrap(~variable)
```

# Variable importance

```{r, fig.height=8}
# Build a combined met-metric dataset
resist %>% 
  select(variable, met, trough_resistance) %>% 
  # We see from the met analysis above that vbdsf and vddsf are highly
  # correlated with nbdsf and nddsf respectively; remove one pair
  left_join(select(met2_wide, -nbdsf, -nddsf), by = "met") ->
  resist_combined

resist_combined %>% 
  pivot_longer(dlwrf:vgrd) %>% 
  ggplot(aes(trough_resistance, value, color = variable)) +
  geom_point() + 
  facet_wrap(variable~name, scales="free_y") + 
  geom_smooth(method = "lm", formula = y ~ x) +
  theme(legend.position = "none", axis.text = element_text(size = 5))

# Fit a Random Forest for each variable (across all treatment severities)
# Note the r.f. functions don't play well with tibbles, so change to data frame
rc_all <- split(as.data.frame(resist_combined), resist_combined$variable)
rf_all <- lapply(rc_all, function(x) 
  randomForest(trough_resistance ~ ., data = x[-1:-2], importance = TRUE)
)

# Importance metrics and partial plot data
import <- list()
partial <- list()
for(var in names(rf_all)) {
  mod <- rf_all[[var]]
  imp <- as.data.frame(importance(mod))
  imp$met_var <- rownames(imp)
  import[[var]] <- as_tibble(imp)

  impvar <- rownames(imp)[order(imp[, 1], decreasing = TRUE)]
  for (i in seq_along(impvar)) {
    pp <- partialPlot(mod, rc_all[[var]], impvar[i], plot = FALSE)
    partial[[paste(var, i)]] <- tibble(var = var, 
                                       met_var = impvar[i],
                                       x = pp$x, y = pp$y)
  }
}

import <- bind_rows(import, .id = "var")
ggplot(import, aes(met_var, `%IncMSE`)) + geom_col() +
  coord_flip() +
  facet_wrap(~var)

partial <- bind_rows(partial)
ggplot(bind_rows(partial), aes(x, y, color = var)) + 
  geom_line() + 
  facet_wrap(~var + met_var, scales = "free")
```