### 2021-05-30 metric_analysis.Rmd
Fix met-metric plots

---
title: "metric analysis"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
    toc_float: yes
    number_sections: true
    code_folding: hide
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)

library(readr)
resistance <- read_csv("ED-outputs/results/metric/resistance_values.csv")
resilience <- read_csv("ED-outputs/results/metric/resilience_values.csv")
met <- read_csv("ED-outputs/results/constant_annual_met_data.csv")
```

# Met data

Take a look at the distribution of met data, and perform a Principal Components
Analysis (PCA) to examine auto-correlation between the various met variables.

```{r, echo=FALSE}
met %>% 
  # hgt is constant; remove
  filter(variable != "hgt") %>%   
  group_by(met, variable) %>% 
  summarise(value = mean(value), .groups = "drop") ->
  met2

library(ggplot2)
theme_set(theme_minimal())
ggplot(met2, aes(met, value)) +
  geom_point() + 
  facet_wrap(~variable, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 90))

library(tidyr)
met2_wide <- pivot_wider(met2, id_cols = "met", names_from = "variable")
pca_met <- prcomp(met2_wide[-1], scale. = TRUE)  # -1 removes met year column 
biplot(pca_met, scale = 0)
pairs(met2_wide[-1])
```

# Metric data

```{r}
ggplot(resistance, aes(trough_resistance, fill = severity)) +
  geom_histogram(bins = 20) + 
  ggtitle("Resistance") +
  facet_wrap(~variable)

ggplot(resilience, aes(resilience, fill = severity)) +
  geom_histogram(bins = 20) + 
  ggtitle("Resilience") +
  facet_wrap(~variable)
```

# Met-metric data

We see from the met analysis above that vbdsf and vddsf are highly
correlated with nbdsf and nddsf respectively; remove one pair.

Take a first look at the relationship between output variables and met variables.

```{r, fig.height=8}
# Build a combined met-metric dataset
resistance %>% 
  select(variable, met, trough_resistance) %>% 
  # We see from the met analysis above that vbdsf and vddsf are highly
  # correlated with nbdsf and nddsf respectively; remove one pair
  left_join(select(met2_wide, -nbdsf, -nddsf), by = "met") ->
  resist_combined

resist_combined %>% 
  pivot_longer(dlwrf:vgrd) %>% 
  ggplot(aes(value, trough_resistance, color = variable)) +
  geom_point() + 
  facet_wrap(variable~name, scales = "free") + 
  geom_smooth(method = "lm", formula = y ~ x, color = "black", linetype = 5) +
  theme(legend.position = "none", axis.text = element_text(size = 5))
```

# Variable importance

Fit a Random Forest model to each output variable (GPP, etc.) and then calculate
variable importance for each model. Make partial dependence plots for the
three most important met variables for each output variable.

```{r, message=FALSE}
# Fit a Random Forest for each variable (across all treatment severities)
# Note the r.f. functions don't play well with tibbles, so change to data frame
library(randomForest)
rc_all <- split(as.data.frame(resist_combined), resist_combined$variable)
rf_all <- lapply(rc_all, function(x) 
  randomForest(trough_resistance ~ ., data = x[-1:-2], importance = TRUE)
)

# Importance metrics and partial plot data
import <- list()
partial <- list()
for(var in names(rf_all)) {
  mod <- rf_all[[var]]
  imp <- as.data.frame(importance(mod))
  imp$met_var <- rownames(imp)
  import[[var]] <- as_tibble(imp)

  impvar <- rownames(imp)[order(imp[, 1], decreasing = TRUE)]
  for (i in seq_along(impvar[1:3])) {
    pp <- partialPlot(mod, rc_all[[var]], impvar[i], plot = FALSE)
    partial[[paste(var, i)]] <- tibble(var = var, 
                                       met_var = impvar[i],
                                       x = pp$x, y = pp$y)
  }
}

import <- bind_rows(import, .id = "var")
library(forcats)
lapply(split(import, import$var), function(x) {
  ggplot(x, aes(fct_reorder(met_var, `%IncMSE`), `%IncMSE`)) + 
    geom_col() +
    coord_flip() +
    xlab("Met variable") +
    ggtitle(unique(x$var))
})

partial <- bind_rows(partial)
ggplot(bind_rows(partial), aes(x, y, color = var)) + 
  geom_line() + 
  xlab("Met var value") + ylab("Metric") +
  ggtitle("Partial dependence plot") +
  facet_grid(var ~ met_var, scales = "free")
```

# Summary

Interestingly, `tmp` (air temperature) and `vbdsf` (sunlight) are among the
top-three most important drivers for all four output variables 
(GPP, NEP, NPP, Rh) for resistance. Atmospheric pressure (`pres`) is
among the top three for both GPP and NPP resistance: the former because of
stomatal effects perhaps (?), and the latter because
of NPP's strong dependence on GPP.