### 2021-05-30 metric_analysis.Rmd
Analyze both resistance and resilience

---
title: "metric analysis"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
    toc_float: yes
    number_sections: true
    code_folding: hide
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(1234)

library(dplyr)

library(readr)
resistance <- read_csv("ED-outputs/results/metric/resistance_values.csv")
resilience <- read_csv("ED-outputs/results/metric/resilience_values.csv")
met <- read_csv("ED-outputs/results/constant_annual_met_data.csv")

resistance %>% 
  select(scn, variable, met, value = trough_resistance, severity) %>% 
  mutate(metric = "resistance") ->
  combined
resilience %>% 
  select(scn, variable, met, value = resilience, severity) %>% 
  mutate(metric = "resilience") %>% 
  bind_rows(combined) ->
  combined
```

# Met data

Take a look at the distribution of met data, and perform a Principal Components
Analysis (PCA) to examine auto-correlation between the various met variables.

```{r, echo=FALSE}
met %>% 
  # hgt is constant; remove
  filter(variable != "hgt") %>%   
  group_by(met, variable) %>% 
  summarise(value = mean(value), .groups = "drop") ->
  met2

library(ggplot2)
theme_set(theme_minimal())
ggplot(met2, aes(met, value)) +
  geom_point() + 
  facet_wrap(~variable, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 90))

library(tidyr)
met2_wide <- pivot_wider(met2, id_cols = "met", names_from = "variable")
pca_met <- prcomp(met2_wide[-1], scale. = TRUE)  # -1 removes met year column 
biplot(pca_met, scale = 0)
pairs(met2_wide[-1])
```

# Metric data

```{r}
ggplot(resistance, aes(trough_resistance, fill = severity)) +
  geom_histogram(bins = 20) + 
  ggtitle("Resistance") +
  facet_wrap(~variable)

ggplot(resilience, aes(resilience, fill = severity)) +
  geom_histogram(bins = 20) + 
  ggtitle("Resilience") +
  facet_wrap(~variable)
```

# Met-metric data {.tabset}

We see from the met analysis above that vbdsf and vddsf are highly
correlated with nbdsf and nddsf respectively; remove one pair.

Take a first look at the relationship between output variables and met variables.

## Resistance

```{r, fig.height=8}
# Build a combined met-metric dataset
combined %>% 
  select(metric, variable, met, value) %>% 
  # We see from the met analysis above that vbdsf and vddsf are highly
  # correlated with nbdsf and nddsf respectively; remove one pair
  left_join(select(met2_wide, -nbdsf, -nddsf), by = "met") ->
  metric_combined

resistance %>% 
  select(variable, met, trough_resistance) %>% 
  left_join(select(met2_wide, -nbdsf, -nddsf), by = "met") ->
  resist_combined

metric_combined %>% 
  pivot_longer(dlwrf:vgrd, values_to = "met_value") ->
  metric_comb_long

p_mm <- ggplot(filter(metric_comb_long, metric == "resistance"), 
            aes(met_value, value, color = variable)) +
  geom_point() + 
  geom_smooth(method = "lm", formula = y ~ x, color = "black", linetype = 5) +
  facet_wrap(~ variable + name, scales = "free") + 
  ggtitle("Resistance") +
  theme(legend.position = "none", axis.text = element_text(size = 5))
print(p_mm)
```

## Resilience

```{r, fig.height=8}
p_mm <- p_mm %+% filter(metric_comb_long, metric == "resilience")
print(p_mm + ggtitle("Resilience"))
```

# Variable importance {.tabset}

Fit a Random Forest model to each output variable (GPP, etc.) and then calculate
variable importance for each model. Make partial dependence plots for the
three most important met variables for each output variable.

```{r, message=FALSE}
# Fit a Random Forest for each variable (across all treatment severities)
# Note the r.f. functions don't play well with tibbles, so change to data frame
library(randomForest)
rc_all <- split(as.data.frame(metric_combined), ~ metric + variable)
rf_all <- lapply(rc_all, function(x) 
  # remove metric, variable, met columns and run RF
  randomForest(value ~ ., data = x[-1:-3], importance = TRUE)
)

# Importance metrics and partial plot data
import <- list()
partial <- list()
for(var in names(rf_all)) {
  mod <- rf_all[[var]]
  metric <- rc_all[[var]]$metric[1]
  variable <- rc_all[[var]]$variable[1]

  imp <- as.data.frame(importance(mod))
  imp$met_var <- rownames(imp)
  imp$metric <- metric
  imp$variable <- variable
  import[[var]] <- as_tibble(imp)

  impvar <- rownames(imp)[order(imp[, 1], decreasing = TRUE)]
  for (i in seq_along(impvar[1:3])) {
    pp <- partialPlot(mod, rc_all[[var]], impvar[i], plot = FALSE)
    partial[[paste(var, i)]] <- tibble(metric = metric,
                                       variable = variable,
                                       met_var = impvar[i],
                                       x = pp$x, y = pp$y)
  }
}

import <- bind_rows(import)
partial <- bind_rows(partial)
```

## Resistance

```{r}
library(forcats)
f_imp_plot <- function(x) {
  ggplot(x, aes(fct_reorder(met_var, `%IncMSE`), `%IncMSE`)) + 
    geom_col() +
    coord_flip() +
    xlab("Met variable") +
    ggtitle(unique(x$variable))
}
lapply(split(filter(import, metric == "resistance"), ~ variable), f_imp_plot)

p_pdp <- ggplot(filter(partial, metric == "resistance"), 
                aes(x, y, color = variable)) + 
  geom_line() + 
  xlab("Met var value") + ylab("Resistance") +
  ggtitle("Partial dependence plot") +
  facet_grid(variable ~ met_var, scales = "free")
print(p_pdp)
```

## Resilience

```{r}
# Re-do above plots but now for resilience
lapply(split(filter(import, metric == "resilience"), ~ variable), f_imp_plot)
p_pdp <- p_pdp %+% filter(partial, metric == "resilience")
print(p_pdp + ylab("Resilience"))
```

# Summary

For **resistance**, `tmp` (air temperature) and `vbdsf` (sunlight) are among the
top-three most important drivers for all four output variables 
(GPP, NEP, NPP, Rh). Higher temperature and more light translate into higher
resistance for all variables.

This is largely true for **resilience** as well, although
`prate` (precipitation) is important for both NPP and Rh; in fact precipitation
is far and away the most important variable for Rh. Here, however,
higher temperature and more light mean _less_ resilience.