### 2021-07-27 C.analysis/paper_analysis.Rmd
Update to manuscript circualted to co-authors at end of July

---
title: "FoRTE Disturbance Paper Analysis"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
    toc_float: yes
    number_sections: true
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

# Objective

## Set up script

```{r setup, include=FALSE, message=FALSE, echo = FALSE, cache = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width = 8, fig.height = 5)
# R 3.6.3
library(dplyr)
library(ggplot2)
library(stringr)
library(stickylabeller)
library(ggpmisc)
library(ggpubr)

BASE_DIR <- "/Users/dorh012/Documents/2020/FoRTE/forte-disturbance"
INPUT_DIR <- file.path(BASE_DIR, "ED-outputs", "results")

# Define the color themes and the graphing aesthetics. 
THEME <- theme_bw()
FORTE_SEVERITY_COLORS <- c("#009E73", "#0072B2", "#D55E00")
names(FORTE_SEVERITY_COLORS) <- c('45 %', '65 %', '85 %')
```

Load the data

```{r}
# Load the ED output data
ED_output <- read.csv(list.files(INPUT_DIR, "exp-constant-yr.csv", full.names = TRUE), stringsAsFactors = FALSE)

# Load the meteorological input, the annual average and the monthly average. 
annual_met <- read.csv(list.files(INPUT_DIR, "constant_annual_met_data.csv", full.names = TRUE), stringsAsFactors = FALSE)
monthly_met <- read.csv(list.files(INPUT_DIR, "constant_annual_met_data-monthly.csv", full.names = TRUE), stringsAsFactors = FALSE)

# Load the results realted to the metrics of ecosystem stability, the log ratio of the time series, resilience, 
# and resistance. 
disturbance <- read.csv(file.path(INPUT_DIR, "metric", "ln_ratio_timeseries.csv"), stringsAsFactors = FALSE) %>% 
  mutate(met = as.character(met))
resilience <- read.csv(file.path(INPUT_DIR, "metric", "resilience_values.csv"), stringsAsFactors = FALSE) %>% 
  mutate(met = as.character(met))
resistance <- read.csv(file.path(INPUT_DIR, "metric", "resistance_values.csv"), stringsAsFactors = FALSE) %>% 
  mutate(met = as.character(met))
```

```{r}
annual_met %>% 
  filter(variable %in% c("ugrd", "vgrd")) %>% 
  mutate(value = value^2) %>% 
  group_by(year, met) %>% 
  summarise(value = sqrt(sum(value))) %>% 
  ungroup() %>%
  mutate(units = "m s-1", 
         description = "total winnd", 
         variable = "grd") -> 
  wind_df

annual_met %>% 
  filter(!variable %in% c("ugrd", "vgrd")) %>% 
  bind_rows(wind_df) %>% 
  filter(variable != "hgt") ->
  annual_met

monthly_met %>% 
  filter(variable %in% c("ugrd", "vgrd")) %>% 
  mutate(value = value^2) %>% 
  group_by(month, mon_time, met) %>% 
  summarise(value = sqrt(sum(value))) %>% 
  ungroup() %>%
  mutate(units = "m s-1", 
         description = "total winnd", 
         variable = "grd") -> 
  wind_df

monthly_met %>% 
  filter(!variable %in% c("ugrd", "vgrd")) %>% 
  bind_rows(wind_df) %>% 
  filter(variable != "hgt") ->
  monthly_met
```

# Baseline Scenario

```{r}
ED_output %>%  
  filter(scn == "harvest_0" & variable  %in% c("GPP", "NPP", "Rh", "NEP", "AGB")) %>% 
  mutate(description = paste0(variable, '\n (', unit,  ')')) %>% 
  dplyr::filter(year == 2020) %>% 
  group_by(variable, unit) %>% 
  summarise(sd = sd(value), mean = mean(value)) %>% 
  ungroup %>% 
  mutate(cv = sd/mean)

```

```{r, fig.width=6.5, fig.height=4, fig.retina = 4}

# # UMBS forest observations taken from table 2 of 
# C.M. Gough, C.S. Vogel, H.P. Schmid, H.-B. Su, P.S. Curtis,
# Multi-year convergence of biometric and meteorological estimates of forest carbon storage,
# Agricultural and Forest Meteorology,
# Volume 148, Issue 2,
# 2008,
# Pages 158-170,
# ISSN 0168-1923,
# https://doi.org/10.1016/j.agrformet.2007.08.004
data.frame(description = c("NPP\n (MgC ha-1 year-1)", "NEP\n (MgC ha-1 year-1)"), 
           mean = c(6.54, 1.53), 
           lower =  c(6.54 - .76, 1.53 - 1.15), 
           upper = c(6.54 + .76,1.53 + 1.15 ), 
           year = c(2001, 2001)) -> 
  obs

ED_output %>%  
  filter(scn == "harvest_0" & variable  %in% c("GPP", "NPP", "Rh", "NEP")) %>% 
  mutate(description = paste0(variable, '\n (', unit,  ')')) -> 
  to_plot

  ggplot() + 
  geom_line(data = to_plot, aes(year, value, group = met), 
            alpha = 0.9, color = "#999999", ) + 
  geom_pointrange(data = obs, aes(x=year, y=mean, ymin=lower, ymax=upper)) + 
  facet_wrap("description", scales = "free", 
             labeller = label_glue('({.l}) {description}')) +
  THEME +
  labs(y = "", x = "Year")  + 
  theme(legend.position = "none") -> 
  plot

plot
ggsave(plot, filename = "figures/baseline.png", width = 6, height = 4.5)

```

# Disturbance Time Series

Log ratio plots show in the Hillbrand manuscript.

```{r}
disturbance %>%  
  mutate(Severity = severity) %>% 
  ggplot() + 
  geom_line(aes(year, value, color = Severity, line = met), alpha = 0.5) + 
  facet_wrap("variable", scales = 'free') +
            # labeller = label_glue('({.l}) {variable}')) +
  THEME +
  scale_color_manual(values = FORTE_SEVERITY_COLORS) + 
  guides(colour = guide_legend("Severity")) + 
  labs(y = "Log Ratio", x = "Years Since Disturbance") -> 
  plot

plot

ggsave(plot, filename = "figures/disturbance.png", width = 6, height = 4.5)

```

# Resistance

```{r}

resistance %>% 
  group_by(variable, severity) %>%  
  summarise(mean = mean(trough_resistance), sd = sd(trough_resistance)) %>% 
  ungroup(.) -> 
  summary_table
```

```{r}
summary_table %>% 
  select(variable, severity, mean) %>% 
  tidyr::spread(severity, mean) %>% 
  mutate(change1 = `65 %` - `45 %` / `45 %`, 
         change2 = `85 %` - `65 %` / `65 %`)

```

```{r, fig.width=5, fig.height=5}
ggplot() + 
  geom_point(data = resistance, aes(severity, trough_resistance, color = severity),
             position = position_jitter(), 
             alpha = 0.75, size = 1.5) + 
    geom_pointrange(data = summary_table, 
                    aes(x=severity, y=mean, ymin=mean-sd, ymax=mean+sd), 
                  color = "black") + 
    ylim(-1.5, 0) +
  facet_wrap("variable",
             labeller = label_glue('({.l}) {variable}')) + 
  labs(y = "Resistance", 
       x = "Severity") + 
  theme_bw(base_size = 10) +
  theme(legend.position = "none") +
  scale_color_manual(values = FORTE_SEVERITY_COLORS) -> 
  plot

ggsave(plot, filename = "figures/resistance.png", width = 6, height = 4.5)
```

# Resilience

```{r}
resilience %>% 
  group_by(variable, severity) %>% 
  summarise(mean = mean(resilience), sd = sd(resilience)) %>% 
  ungroup(.) -> 
  summary_table
```

```{r}
summary_table %>% 
  select(variable, severity, mean) %>% 
  tidyr::spread(severity, mean) %>% 
  mutate(change1 = `65 %` - `45 %` / `45 %`, 
         change2 = `85 %` - `65 %` / `65 %`)

```

```{r, fig.width=5, fig.height=5}
ggplot() + 
  geom_point(data = resilience, aes(severity, resilience, color = severity),              
             position = position_jitter(), 
             alpha = 0.75, size = 1.5) + 
  facet_wrap("variable", labeller = label_glue('({.l}) {variable}')) + 
  geom_pointrange(data = summary_table, 
                    aes(x=severity, y=mean, ymin=mean-sd, ymax=mean+sd), 
                    color = "black") + 
  labs(y = "Resilience", 
        x = "Severity") + 
  ylim(0, 0.035)+
  theme_bw(base_size = 10) +
  theme(legend.position = "none") +
  scale_color_manual(values = FORTE_SEVERITY_COLORS) -> 
  plot

  ggsave(plot, filename = "figures/resilience.png", width = 6, height = 4.5)
```

# Resistance vs Resilience

```{r, fig.width=5, fig.height=5}
resilience %>% 
  left_join(resistance) %>%  
  ggplot(aes(trough_resistance, resilience, color = severity)) + 
  geom_point(alpha = 0.75, size = 1.5) + 
  facet_wrap("variable")+ #scales = "free",  
            # labeller = label_glue('({.l}) {variable}')) + 
  labs(x = "Resistance", 
       y = "Resilience", 
       title = "Resilience vs. Resistance") + 
  theme_bw(base_size = 10) +
  scale_color_manual(values = FORTE_SEVERITY_COLORS) + 
  guides(colour = guide_legend("Severity")) -> 
  plot

ggsave(plot, filename = "figures/resistance_resilience.png", width = 6, height = 4.5)

```

```{r, fig.width=5, fig.height=5}
my.formula <- y ~ x

resilience %>% 
  left_join(resistance) %>%  
  ggplot(aes(trough_resistance, resilience, color = severity)) + 
  geom_point(alpha = 0.75, size = 1.5) + 
  geom_smooth(method = "lm", size = 0.5, alpha = 0.2) + 
  stat_cor(method = "pearson", 
           aes(label = ..r.label..),
           size = 3, 
           label.y.npc = "bottom",
           r.accuracy = 0.01) + 
  labs(x = "Resistance", 
       y = "Resilience", 
       title = "Resilience vs. Resistance") + 
  theme_bw(base_size = 10) +
  scale_color_manual(values = FORTE_SEVERITY_COLORS) + 
    facet_wrap("variable", 
               scales = "free", 
               labeller = label_glue('({.l}) {variable}')) + 
  guides(colour = guide_legend("Severity")) -> 
  plot
  ggsave(plot, filename = "figures/resistance_resilience3.png", width = 6, height = 4.5)

```

```{r}

resilience %>% 
  left_join(resistance) %>%  
  ggplot(aes(trough_resistance, resilience, color = severity)) + 
  geom_point(alpha = 0.75, size = 1.5) + 
  geom_smooth(method = "lm", size = 0.5, alpha = 0.2) + 
  stat_cor(method = "pearson", aes(label = ..r.label..),
           size = 3, label.x.npc = "left", label.y.npc = "bottom", 
           r.accuracy = 0.01) + 
  labs(x = "Resistance", 
       y = "Resilience", 
       title = "Resilience vs. Resistance") + 
  theme_bw(base_size = 10) +
  scale_color_manual(values = FORTE_SEVERITY_COLORS) + 
  guides(colour = guide_legend("Severity"))

```

# Investigating the relationship between met and the metrics

Annual met values (by design they should be constant/flat over the entire time period)

```{r, fig.width=5, fig.height=5}
annual_met %>% 
  mutate(description = paste0(description, ' (', units, ')')) %>% 
  ggplot(aes(year, value, color = met, group = met)) + 
  geom_line() +
   facet_wrap("description", scales = 'free', strip.position = "top",
             labeller = label_wrap_gen(width = 20, multi_line = TRUE)) +
  THEME +
  labs(y = NULL, x  = NULL,  title = "Annual Met") + 
  theme(legend.position = "none")
```

Monthly values, take a look at the pattern that is going on within each year.

```{r, fig.width=10, fig.height=5.5}
monthly_met %>% 
  mutate(units = if_else(variable == "vgrd", 'm s-1', units)) %>% 
  mutate(description = paste0(str_to_title(description), '\n(', units, ')')) -> 
  data

data %>% 
  filter(met == "2019") -> 
  data_2019

ggplot() + 
  geom_line(data = data, aes(mon_time, value, group = met, line = met), 
            alpha = 0.8, color = "#999999") +
  geom_line(data = data_2019, aes(mon_time, value, color = met, group = met, line = met), 
            color = "#D55E00", size = 1) +
   facet_wrap("description", scales = 'free', strip.position = "top",
             labeller = label_wrap_gen(width = 20), ncol = 2) +
  labs(y = NULL, x  = NULL) +
  THEME +
  scale_x_continuous(labels = c("Jan", "Jun", "Dec"), breaks = c(1, 6, 12)) +
  theme(legend.position = "none") -> 
  plot

plot

ggsave(plot, filename = "figures/monthly_met.png", width = 6, height = 10)
```

```{r}
annual_met %>%  
  mutate(met = as.integer(met)) %>% 
  select(met, met_var = variable, met_value = value, met_des = description) %>% 
  distinct() -> 
  met_data

resilience %>%
  left_join(resistance) %>%  
  mutate(met = as.integer(met)) %>% 
  left_join(met_data) %>%  
  select(scn, severity, met, variable, resilience, resistance  = trough_resistance, met_var, met_value, met_des) -> 
  metrics_v_met
```

## Met vs. Metrics  {.tabset}

```{r}
# Save a copy of the meteorological variables 
vars <- unique(metrics_v_met$met_var)

# Create the plots of met vs resistance. 
lapply(vars, function(met_var_name){

  des <- unique(metrics_v_met[metrics_v_met$met_var == met_var_name, ]$met_des)

  metrics_v_met %>% 
    filter(met_var == met_var_name) %>%  
    ggplot(aes(met_value, resistance, color = severity)) + 
    geom_point() + 
    facet_wrap('variable', scales = "free", nrow = 1) + 
    THEME + 
    scale_color_manual(values = FORTE_SEVERITY_COLORS) +
    labs(x  = des, title = 'Resistance') -> 
    out

  return(out)

}) ->
  resistance_plots
names(resistance_plots) <- vars

# Create the plots of met vs resilience 
lapply(vars, function(met_var_name){

  des <- unique(metrics_v_met[metrics_v_met$met_var == met_var_name, ]$met_des)

  metrics_v_met %>% 
    filter(met_var == met_var_name) %>%  
    ggplot(aes(met_value, resilience, color = severity)) + 
    geom_point() + 
    facet_wrap('variable', scales = "free", nrow = 1) + 
    THEME + 
    scale_color_manual(values = FORTE_SEVERITY_COLORS) +
    labs(x  = des, title = 'Resilience') -> 
    out

  return(out)

}) ->
  resilience_plots
names(resilience_plots) <- vars
```

### downward long wave radiation

```{r, fig.width=10, fig.height=4}
resistance_plots$dlwrf
```

```{r, fig.width=10, fig.height=4}
resilience_plots$dlwrf
```

### near infrared beam downward solar radiation

```{r, fig.width=10, fig.height=4}
resistance_plots$nbdsf
```

```{r, fig.width=10, fig.height=4}
resilience_plots$nbdsf
```

### near IR diffuse downward solar radiation

```{r, fig.width=10, fig.height=4}
resistance_plots$nddsf
```

```{r, fig.width=10, fig.height=4}
resilience_plots$nddsf
```

### visible beam downward solar radiation

```{r, fig.width=10, fig.height=4}
resistance_plots$vbdsf
```

```{r, fig.width=10, fig.height=4}
resilience_plots$vbdsf
```

### visible diffuse downward solar radiation

```{r, fig.width=10, fig.height=4}
resistance_plots$vddsf
```

```{r, fig.width=10, fig.height=4}
resilience_plots$vddsf
```

### precipitation rate

```{r, fig.width=10, fig.height=4}
resistance_plots$prate
```

```{r, fig.width=10, fig.height=4}
resilience_plots$prate
```

### atmospheric pressure

```{r, fig.width=10, fig.height=4}
resistance_plots$pres
```

```{r, fig.width=10, fig.height=4}
resilience_plots$pres
```

### geopotential height

```{r, fig.width=10, fig.height=4}
resistance_plots$hgt
```

```{r, fig.width=10, fig.height=4}
resilience_plots$hgt
```

### zonal wind

```{r, fig.width=10, fig.height=4}
resistance_plots$ugrd
```

```{r, fig.width=10, fig.height=4}
resilience_plots$ugrd
```

### meridional wind

```{r, fig.width=10, fig.height=4}
resistance_plots$vgrd
```

```{r, fig.width=10, fig.height=4}
resilience_plots$vgrd
```

### specific humidity

```{r, fig.width=10, fig.height=4}
resistance_plots$sh
```

```{r, fig.width=10, fig.height=4}
resilience_plots$sh
```

### air temperature

```{r, fig.width=10, fig.height=4}
resistance_plots$tmp
```

```{r, fig.width=10, fig.height=4}
resilience_plots$tmp
```

## Met vs Metrics observations

Breifly looking through the different pannels what I've notcies is that patterns between meterology values and the stability meteric are more linear looking for resistance comapred to resillience.

The three meterology variables that are the most interesting visually are

* air temperature 
* precipitation rate
* air temperature

# TO DO

Kalyn do you want to take a stab that a random variancne